opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-05-11 chapters/00_abstract.tex
Added todos in introduction
Commited updates from Henrik/Pascal

Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
Despite its widespread application, the analysis of the diffractograms still presents a significant bottleneck
in high-throughput experimentation that remains to be solved in order to match the increased throughput
of the prior stages of experimentation.

In the current landscape of powder X-ray diffraction analysis, machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction analysis.
A notable difficulty in applying machine learning to this domain is the lack of labeled datasets, i.e.\
diffraction patterns originating from experiment with accompanying structure solutions, which has relegated machine learning
researchers to train on simulated data.
Since these simulations largely fail to accurately reflect the experiment, the performance of models trained
on this simulated data by and large fails to transfer to experiment and provide value in practice.
We aim to remedy this by providing a freely available and easily accessible dataset of partially labeled
powder diffraction data, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
We hope that this can guide machine learning research toward more realistic handling of experimental imperfections to enable the improvement of current automated analysis workflows.

>>>>

Powder X-ray diffraction (pXRD) experiments are a cornerstone for materials structure characterization.
Despite their widespread application, the analysis of pXRD diffractograms still presents a significant automation challenge and thus a bottleneck in high-throughput experimentation that remains to be solved in order to match increasing throughput of automated materials synthesis in (self-driving) labs.
In the current landscape of powder X-ray diffraction analysis, machine learning has emerged as a promising research direction to resolve this bottleneck by enabling automated powder diffraction analysis.
A notable difficulty in applying machine learning to this domain is the lack of sufficiently sized experimental datasets, which has relegated machine learning researchers to train primarily on simulated data. Since simulations largely fail to accurately reflect the experiment, the performance of models trained on only simulated data often lacks transferability to experimental data and thus fails to provide value in practice.
With the Open Experimental Powder X-Ray Diffraction Database (opXRD), we aim to remedy this by providing an openly available and easily accessible dataset of partially labeled experimental powder diffractograms, providing machine learning researchers with a large quantity of real experimental diffractograms collected from a broad range of samples.
We provide almost barrier-free software tools that allow experimental researchers to find their data and make it accessible - in virtually any widely used format.
We collected XXX labeled and YYY unlabeled diffractograms from a wide spectrum of materials classes, which establishes the first version of the opXRD database.
We hope that this ongoing effort can guide machine learning research toward fully automated analysis of pXRD data and thus enable future self-driving materials labs.

### 2024-05-11 chapters/01_introduction.tex
Commited updates from Henrik/Pascal

\begin{itemize}
    \item Plan for Figures / Tables
    \begin{itemize}
        \item Figure 1: Multiple panels showing histograms of some key metrics: Space group, unit cell volume (normalizable?), ...
        \item Table 1: List of existing powder XRD datasets
        \item Table 2: List of existing ML approaches for pXRD
        \item Table 3: New pXRD datasets offered by this paper
    \end{itemize}

    \item Additional thoughts
    \begin{itemize}
        \item General structure of the paper:
        \begin{itemize}
            \item Introduction
            \item Introduce existing datasets and already show why they are not enough / too small
            \item Summarize SOTA ML that has been applied to experimental pXRD, discuss the limits and gap between performance on simulated data and exp. data (find some good example papers where this is very apparent)
            \item Summary of our dataset, including paragraphs of the different contributions by other groups; comparison with existing datasets, especially in terms of size
        \end{itemize} 
    \end{itemize}
\end{itemize}

@@

The advent of high-throughput experiments, holds the prospect of accelerating the speed materials discovery multiple times over.
The synthesis and probing of novel materials is becoming increasingly efficient and automated,
which has correspondingly increased the throughput of samples in these steps of the experimentation pipeline.

@@

% On Rietveld refinement
The objective of powder x-ray diffraction experiments is to extract as much information about the crystal structure
of the sample as possible, or if the sample is multi-phase, also identify the phases present in the sample and quantify
their fractions.
This is done through a technique called Rietveld refinement: A working model of the crystal structure, starting
from an initial guess provided by the operator of the analysis software, is assumed and the corresponding diffractogram
is simulated.
The parameters of the working model of the sample are then adjusted to minimize the difference between the simulated
diffraction pattern and the pattern observed in experiment.
Critically, this process does not generally guarantee to guide the working model towards a global minimum of the difference
in diffractograms and hence arrive at an accurate model of the sample.
In general one can only assume the refinement process to guide the working model towards the nearest local minimum
starting from the initial guess.
Hence, the result of the refinement procedure is in general only as good as the initial guess it was provided with.
Still this workflow often works well, since usually the operator of the experiment will be able
to provide an at least somewhat accurate initial guess of the crystal structure.\\

However, employing manually operated Rietveld Refinement is not without issues.
It is certainly not scalable to the degree that would be required to keep up with advances
in throughput and efficiency in other steps of the experimentation pipeline, even just due to the manual effort involved.
But beyond the lacking scalability of this workflow we also wish to highlight the following issue:
Providing the analysis software with an initial guess of the structure requires manually or algorithmically
consulting a reference library of crystal structures with corresponding simulated or sometimes real diffractograms.
From that reference library a crystal structure must be identified that is likely to be similar
to that of the investigated sample which can then serve as the initial guess provided to the analysis software.
Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
The best that can be done to remedy this is to continually update the library both on the side
of the distributor and the user, which is an undesirable overhead.
Especially for novel material discovery, it may be impossible to find a sufficiently accurate match in the reference
library to apply Rietveld Refinement successfully.

Machine learning is a very promising candidate to eliminate both the need for manual inputs and the need for reference
libraries in powder x-ray diffraction analysis, resulting in a streamlined and automated analysis process that can
keep pace with an automated high-throughput experimentation environment. \\

Models can be trained to predict crystal structure properties given only the diffractogram itself, given that
the model can be trained on dataset consisting of pairs of diffractograms and corresponding crystal structures.
So far however, due to an absence of sufficiently large labeled datasets with diffractograms from experiment,
machine learning in this domain has largely relied on simulated diffractograms.
Note, that reference libraries which are used for Rietveld Refinement by in large use simulated diffractograms
rather than diffractograms from experiment and hence cannot either provide a dataset with diffractograms coming
from experiment. \\

Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters, spacegroup and crystallite size from simulated diffractograms but the performance drops off immensely when these models are applied to data originating from experiment.

The patterns observed in experiment suffer from a host of imperfections that may result in a partial loss of information or make extracting the structural information more difficult. \\
However, there is still reason to believe that the strong performance observed in the domain of purely simulated diffractograms can to a large extent translate to predicting crystal structure properties from diffractograms originating from experiment, given that models are trained on a dataset that accurately reflects diffractograms as they appear in experiment.

However due to the scarcity of labeled diffractograms originating from experiment, collecting a labeled dataset
of real experimental powder x-ray diffraction data of sufficient size to train a model
that can be useful in practice is next to impossible and is not the aim of this dataset.
Rather we hope that this dataset can enable machine learning researchers to evaluate how closely their simulations
represent actual experimental data, identify what features are missing and modify their simulations algorithms accordingly to arrive at a simulation that nearly perfectly mirrors how data appears in reality.\\

There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern. This would for example necessitate modelling absorption of the incoming X-ray radiation and its re-emission in the form of characteristic radiation and modeling the effect of stress and motion within the crystal on the diffractogram. \\
An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset. \\
As an example of this could be done we can look at results from other domains, where a transfer of core data between different "styles" of representation was attempted:
%\cite{Ganin2015} -> Attempted to

@@

For a set of given sample properties there is only one corresponding simulated diffractogram, but there may an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
, sample impurities and lattice defects.
In order for a simulated dataset to closely resemble experiment it is critical that the simulatio also represents these
statistical variations.
Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
originated from experiment and cannot be accomplished by simply training a classifier.
Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
simulation as well.\\

Secondly we would like to also present this dataset as a challenge: The fraction of labeled data can be used to easily numerically evalute the performance of the model on a broad range of data as it appears in experiment, which can be used to gauge how well models would perform in practice. \\
But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure. From the predicted crystal structure, the corresponding diffractogram can be simulated, which can then be compared to the real diffractogram that was used as input for the prediction. While a nearly matching diffractogram doesn't necessarily mean a correctly predicted structure, a non-matching diffractogram certainly means an incorrectly predicted structure. \\

Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
to some of the practical difficulties that need to be addressed in applying a model to a wide variety of x-ray diffractograms
collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
which is something that any model that aims to perform well on this dataset must be equipped to handle.

>>>>

@@

\begin{figure*}[!ht]
    \centering
    \includegraphics{figures/overview.pdf}
    \caption{Overview of the data collection pipeline. Datasets are submitted using an online submission form or our submission helper software. After post-processing and data homogenization, we create a Zenodo entry for each user submission and subsequently include the submission in the opXRD database.}
    \label{fig:overview}
\end{figure*}

The advent of high-throughput experiments holds the prospect of significantly accelerating the speed of materials discovery.
The synthesis and characterization of novel materials is becoming increasingly efficient and automated,
which increases the throughput of samples in experimentation pipelines.

@@

% On Rietveld refinement
After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
The most important method to determine the crystal structure of a new material is powder X-ray diffraction. 
%In the case of multi-phase samples, this includes identifying the phases present in the sample and quantifying their fractions.
To analyze pXRD data, an initial guess is optimized using Rietveld refinement. In this process, an initial crystal structure, including unit cell parameters, provided by the operator of the analysis software, is fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.
%The unit cell parameters of the crystal structure are adjusted to minimize the difference between the simulated diffraction pattern and the pattern observed in the experiment. 
As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.

Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
Furthermore, initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database.
%But beyond the lack of scalability of this workflow we also wish to highlight the following issue:
%Providing the analysis software with an initial guess of the structure requires manually or algorithmically
%consulting a reference library of crystal structures with corresponding simulated or sometimes real diffractograms.
%From that reference library, a crystal structure must be identified that is likely to be similar
%to that of the investigated sample which can then serve as the initial guess provided to the analysis software.
%Since the number of chemically viable crystal structures is infinite, any such library will always be incomplete.
However, in the case of novel observed structure types, simply matching to a database of known structures will not yield good results. Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
%The best that can be done to remedy this is to continually update the library both on the side
%of the distributor and the user, which is an undesirable overhead.
%Especially for novel material discovery, it may be impossible to find a sufficiently accurate match in the reference
%library to apply Rietveld Refinement successfully.

Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment.
%eliminate both the need for manual inputs and the need for reference
%libraries in powder x-ray diffraction analysis, resulting in a streamlined and automated analysis process that can
%keep pace with an automated high-throughput experimentation environment.
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement.
%However, the current limiting factor is the availability of sufficiently large labeled datasets.
So far, however, due to an absence of labeled datasets with experimental diffractograms, machine learning in this domain has largely relied on diffractograms simulated from known structures (such as from the ICSD) or, most recently, from generated synthetic crystals.

Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters, spacegroup, and crystallite size from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

At the moment, the number of openly available datasets for experimental powder data is very limited.
The main sources for experimental patterns are XXX, YYY, and ZZZ.
In this publication, we provide a new dataset that collects a broad range of patterns from experiment, exceeding the currently available amount of openly accessible experimental patterns by XX.
The dataset has been curated by collecting the accumulated powder data from multiple large research groups and institutions with high-throughput XRD facilities.

% How to contribute? This aspect deserves a separate paragraph
We further provide an online platform that can be used to easily submit new pXRD data to the dataset, which will be used to extend the existing dataset in the future (Figure~\ref{fig:overview}).
TODO: Make this longer?

Collecting a labeled dataset of raw experimental powder X-ray diffraction data of sufficient size to directly train a model that predicts crystal structure information is still out of reach.
The most advanced ML models currently are trained on approximately x simulated diffractograms, which is x orders of magnitude higher than the initial state of the opXRD database presented here.
However, our dataset still holds significant value in practice. First of all, the dataset can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.

Furthermore, we hope that our dataset can enable machine learning researchers to evaluate how closely their simulations
represent actual experimental data, identify what features are missing, and modify their simulation algorithms accordingly to arrive at a simulation that mirrors how data appears in reality.
One can bring the simulated patterns closer to experiment by improving the simulation itself by properly modeling more of the observed physical effects such as preferred orientation, peak shapes resulting from different sources, etc.
Additionally, some of the observed imperfections, such as background and noise, can be modeled using machine learning approaches.
This is similar to the concept of style transfer\cite{Gatys2016,Ganin2015} found in the domain of machine learning applied to images.
Here, the objective is to obtain a classification or regression model when only labeled images of one style (simulated data) and unlabeled or partially labeled images of another style (experimental data) are available.
The broad range of available experimental samples contained in our dataset makes it possible to apply these state-of-the-art ML approaches to the domain of powder XRD analysis.
We hope that this will yield more advanced automated ML approaches in the future, which can readily be applied in high-throughput experimentation pipelines.

%We hope that this will yield more attempts of ML-guided XRD analysis in the future, where also open challenges such as different angle step sizes, different radiation sources, will be addressed.

%However, there is still reason to believe that the strong performance observed in the domain of purely simulated diffractograms can to a large extent translate to predicting crystal structure properties from diffractograms originating from experiment, given that models are trained on a dataset that accurately reflects diffractograms as they appear in experiment.

%There are roughly speaking two paths towards improving the simulation: The first is to more accurately model the sample and the processes taking place within it during the simulation of the powder diffraction pattern. This would for example necessitate modelling absorption of the incoming X-ray radiation and its re-emission in the form of characteristic radiation and modeling the effect of stress and motion within the crystal on the diffractogram.

%An alternative route to obtaining a more accurate simulation is to try to learn these "imperfections" compared to an idealized simulation using machine learning techniques by extracting them from this dataset.

%As an example of this could be done, we can investigate results from other areas of research, where a transfer of data content between different "styles" of representation was attempted:

%Ganin et al.\cite{Ganin2015} proposed a mechanism for training a classifier which performs well independently of which style the input data is represented in, which only requires labeled data from one style and unlabeled data from other styles. This corresponds precisely to the situation at hand, where the number of labeled simulated diffractograms is virtually unlimited and a large amount of unlabeled real diffractograms are available, but labeled real diffractogram are in short suppply. This algorithm could be applied to train a "style-invariant" structure properties prediction model using a combination of our dataset and simulated labeled patterns.

%Another approach by Gatys et al. \cite{Gatys2016} achieves style transfer of images, by requiring that the content of the output image remain the same, while the style must match that of another image, chosen to represent the target style. This procedure could be applied to transfer the style of simulated diffractorams to that of real diffractograms, which could then be used to train the model data that is more faithful to powder diffractograms as they appear in experiment.

@@

%For a set of given sample properties there is only one corresponding simulated diffractogram, but there may be an infinite amount of variations in how the diffractogram of a sample could look like in experiment for example due varying measurement apperatus, noise
%, sample impurities and lattice defects.
%In order for a simulated dataset to closely resemble experiment it is critical that the simulation also represents these
%statistical variations.
%Determining the nature of these variations is a much a harder task than making individual diffractograms look like they
%originated from experiment and cannot be accomplished by simply training a classifier.
%Nevertheless, we hope that this dataset can at least provide clues of how to take this aspect into account during
%simulation as well.

%But even the unlabeled portion of the dataset can be used to test the performance of models, if the model attempts to predict the entire crystal structure. From the predicted crystal structure, the corresponding diffractogram can be simulated, which can then be compared to the real diffractogram that was used as input for the prediction. While a nearly matching diffractogram doesn't necessarily mean a correctly predicted structure, a non-matching diffractogram certainly means an incorrectly predicted structure.

%Finally, by inviting the application of models to data as it appears in experiment we would like to bring attention
%to some of the practical difficulties that need to be addressed in applying a model to a wide variety of X-ray diffractograms
%collected from different sources: For example the angle and the step size of the diffractograms will differ between sources, 
%which is something that any model that aims to perform well on this dataset must be equipped to handle.

### 2024-05-11 chapters/02_existing_datasets.tex
Added todos in introduction
Commited updates from Henrik/Pascal

% Source of png : https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement

@@

% -> PDF (https://www.icdd.com/) -> 
%- 1200k "material data sets" (structures?) 
% + 19 k entries with "Experimental raw data digital patterns"; 
%  Proprietary
%The PDF itself lists the following data sources:

\begin{itemize}
\item Powder Diffraction File (ICDD)
\begin{itemize}
    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
\end{itemize}
\end{itemize}

% ->  COD (https://www.crystallography.net/cod/) -
% -  513k structures,
% - Fully open

>>>>

@@

% -> PDF (https://www.icdd.com/) -> 
%- 1200k "material data sets" (structures?) 
% + 19 k entries with "Experimental raw data digital patterns"; 
%  Proprietary (% PDF pricing: https://www.icdd.com/assets/files/2023-Purchase-Options.pdf)
%The PDF itself lists the following data sources:

\paragraph{Existing datasets} TODO: Summarize the existing datasets shortly, including number of entries, etc.

\begin{itemize}
\item RRUFF
\item Powder Diffraction File (ICDD)
%\begin{itemize}
%    \item ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA
%    \item ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database
%    \item NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA
%    \item MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)
%    \item ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA
%    \item TODO: Include how many of those are actually experimental pXRD! If we cannot find this information, we need to ask them directly.
%\end{itemize}
\item Scattered experimental datasets on Zenodo (TODO: roughly estimate size)
\item ...
\end{itemize}

% ->  COD (https://www.crystallography.net/cod/)
% -  513k structures,
% - Fully open

### 2024-05-11 chapters/04_our_dataset.tex
Commited updates from Henrik/Pascal

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information. \\
Currently, institutions that contributed towards our dataset include:
\begin{itemize}
    \item Institute of Nanotechnology at Karlsruhe institute of technology
    \item University of Southern California
    \item Lawrence Berkeley National Laboratory
\end{itemize}

We are still in the process of growing the dataset and additional contributions are still very welcome. 
To find out more about how to contribute to this dataset, visit our website specially designed for the purpose of collecting this dataset, https://xrd.aimat.science \\

A major advantage of this dataset over other comparable datasets is that is is very easy to handle in python and in PyTorch.
Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data\_dirpath). The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method. \\
The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.

>>>>

\begin{figure*}[!ht]
    \centering
    \missingfigure{} 
    \caption{Statistics, histograms, etc. of our dataset.}
    \label{fig:statistics}
\end{figure*}

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information. \\
Currently, institutions that contributed to our dataset include:
\begin{itemize}
    \item Institute of Nanotechnology at Karlsruhe Institute of Technology
    \item University of Southern California
    \item Lawrence Berkeley National Laboratory
\end{itemize}

We are still in the process of growing the dataset and additional contributions are still very welcome. To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science \\

### 2024-05-11 chapters/05_how_to_use.tex
Commited updates from Henrik/Pascal

% Notes:
% The subsections of this section should be the 
% individual contributions that we get from labs

% Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiment, some of them labeled with corresponding structural information. \\
Currently, institutions that contributed towards our dataset include:
\item Institute of Nanotechnology at Karlsruhe institute of technology
\item University of Southern California
\item Lawrence Berkeley National Laboratory

We are still in the process of growing the dataset and additional contributions are still very welcome. 
To find out more about how to contribute to this dataset, visit our website specially designed for the purpose of collecting this dataset, https://xrd.aimat.science \\

A major advantage of this dataset over other comparable datasets is that is is very easy to handle in python and in PyTorch. \\
Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data_dirpath). The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method. \\
The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.

>>>>

TODO: A small tutorial with a few python commands to show how easy it is to use our dataset through the provided API.

A major advantage of this dataset over other comparable datasets is that it is very easy to handle in python and in PyTorch.
Our accompanying python library xrdpattern (https://github.com/aimat-lab/xrdpattern) provides a means to read in as much of the dataset as fits in memory simply by using PatternDB.load(data\_dirpath). The patterns attribute of this class is a list of the indivdual pattern diffractograms, each of which supports a standardize, plot and to tensor method. \\
The standardize method returns a "standardized" version of the pattern with a fixed angle range, a fixed number of entries and intensities normalized to the $[0,1]$ interval.
